---
title: "Explore Two Variables in R"
output: html_notebook
---

## Load diamond data set to explore
```{r}
library(ggplot2)
diam <- diamonds
```

## create a scatterplot of price vs x.
```{r}
ggplot(data = diam, aes( x = x, y = price)) +
  geom_point()
```
```{r}
with(diam,cor.test(price,x,method = 'pearson'))  
with(diam,cor.test(price,y,method = 'pearson'))  
with(diam,cor.test(price,z,method = 'pearson'))  
```
## scatter plot of price vs depth
```{r}
 ggplot(data = diam, aes( x = depth, y = price)) +
  geom_point(alpha = 1/100) +
  scale_x_continuous(breaks = seq(0,80,2))
```

## Scatterplot of price vs carat and omit the top 1% of price and carat

```{r}
ggplot(data = diam, aes( x = carat, y = price)) +
  geom_point() +
  xlim (0,quantile(diam$carat,0.99)) +
  ylim (0,quantile(diam$price,0.99))
```

## Create a scatterplot of price vs. volume (x * y * z)

```{r}
diam$volume <- (diam$x * diam$y * diam$z)
ggplot(data = diam, aes( x = volume, y = price)) +
  geom_point()

#diamond with 0 volume
library(plyr)
library(dplyr)
count(filter(diam,volume==0)) 

with(filter(diam,!(volume == 0|volume >=800)), cor.test(price,volume))
```
# Subset the data to exclude diamonds with a volume
# greater than or equal to 800. Also, exclude diamonds
# with a volume of 0. Adjust the transparency of the
# points and add a linear model to the plot.

```{r}
ggplot(data = filter(diam,!(volume == 0|volume >=800)), aes( x = volume, y = price)) +
  geom_point(alpha= 1/100) +
  geom_smooth(method = 'lm')
```

# create a new data frame containing diamonds by clarity
#       (1) mean_price
#       (2) median_price
#       (3) min_price
#       (4) max_price
#       (5) n

```{r}
diamondsByClarity <- diamonds %>%
  group_by(clarity) %>%
  summarise(mean_price = mean(price),
            median_price = median(price),
            min_price = min(price),
            max_price = max(price),
            n = n())
```

```{r}
diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))
```

```{r}
p1 <- ggplot(data = diamonds_mp_by_clarity, aes(x = clarity,y = mean_price, fill = clarity)) +
  geom_bar(stat = "identity")

p2 <- ggplot(data = diamonds_mp_by_color, aes(x = color,y = mean_price, fill = color)) +
  geom_bar(stat = "identity")

library(gridExtra)
grid.arrange(p1,p2,nrow = 1)
```













