---
title: Predict survival on the Titanic 
output:
  html_notebook: default
---

```{r,echo=FALSE}
library(dplyr)
```

#### Import Training and Test dataset
```{r}
titanic.train <- read.csv("train.csv")
titanic.test <- read.csv(("test.csv"))
```

```{r}
test <- titanic.test
test$Survived <- NA
test1 <- test %>% select(PassengerId,Survived,Pclass:Embarked)
titanic.full <- rbind(titanic.train,test1)
rm(test1)
rm(test)
```


#### Explore and prepare data for analysis
```{r}
summary(titanic.full)
```
*Replace NA with mean value for fare*
```{r}
titanic.full$Fare <- ifelse(is.na(titanic.full$Fare),
                            mean(titanic.full$Fare,na.rm = TRUE),
                            titanic.full$Fare)
```
*Replacing Age Missing value with the median Age*
```{r}
titanic.full$Age <- ifelse(
  is.na(titanic.full$Age), 
  median(titanic.full$Age,na.rm = TRUE),
  titanic.full$Age
  )
```
*Adding Family Indicator and Cabin Type*
```{r}
titanic.full$SibSp <- ifelse(titanic.full$SibSp > 0, 1,0)
titanic.full$Parch <- ifelse(titanic.full$Parch > 0, 1,0)
titanic.full$Family <- ifelse((titanic.full$Parch + titanic.full$SibSp) > 0,1,0)
titanic.full <- subset(titanic.full,select = -c(SibSp,Parch,Name))
titanic.full$Cabin <-substr(titanic.full$Cabin,1,1)
titanic.full$Cabin <-ifelse(titanic.full$Cabin =='',"X",titanic.full$Cabin)  # No Cabin (X)
titanic.full$Embarked <-as.factor(ifelse(as.character(titanic.full$Embarked) =="","S",as.character(titanic.full$Embarked)))
```

*Move data back to training and test set*
```{r}
titanic.train.clean <- titanic.full[1:891,]
titanic.test.clean <- titanic.full[892:1309,]
titanic.test.clean <- titanic.test.clean[,!(names(titanic.test.clean) %in% c("Survived"))]
```

####Exploratory Data Analysis and Plots
*Servived passanger proportion*
```{r}
round(prop.table(table(titanic.train.clean$Survived)),2) # 38% servived
```
*Gender wise passanger proportion*  
```{r}
round(prop.table(table(titanic.train.clean$Sex)),2)  # 35/65
round(prop.table(table(titanic.train.clean$Survived,titanic.train.clean$Sex)),2)
library(ggplot2)
ggplot(data = titanic.train.clean,aes(x = Age)) +
  geom_histogram(aes(fill = as.factor(Survived)),binwidth = 2) +
  facet_wrap(~ Sex)
```
Female of all age was able to survived better, male did not made it.

*PClass wise passanger proportion*
```{r}
round(prop.table(table(titanic.train.clean$Pclass)),2)
round(prop.table(table(titanic.train.clean$Survived,titanic.train.clean$Pclass)),2)

ggplot(data = titanic.train.clean,aes(x = Age)) +
  geom_histogram(aes(fill = as.factor(Survived)),binwidth = 2) +
  facet_wrap(~ Pclass)
```
Plot indicates PClass -1 pasanger survival was better among other PClasses. Survival chance in Pclass 2 was 50%. There was only 13% passengers survivived in PClass 3 even through this class covering the 55% passengers. 

*Cabin wise passanger proportion*
```{r}
round(prop.table(table(titanic.train.clean$Cabin)),2)
round(prop.table(table(titanic.train.clean$Survived,titanic.train.clean$Cabin)),2)

```
Survival percent of passangers in cabin B,C,D,E is better compared to A and F. Cabin X which is default cabin affected most. 

*Embarked wise passanger proportion*
```{r}
round(prop.table(table(titanic.train.clean$Embarked)),2)
round(prop.table(table(titanic.train.clean$Survived,titanic.train.clean$Embarked)),2)

ggplot(data = titanic.train.clean,aes(x = Age)) +
  geom_histogram(aes(fill = as.factor(Survived)),binwidth = 2) +
  facet_wrap(~ Embarked)
```
Passangers from Embarked S is affected most.

*Family wise passanger proportion*
```{r}
round(prop.table(table(titanic.train.clean$Family)),2)
round(prop.table(table(titanic.train.clean$Survived,titanic.train.clean$Family)),2)

ggplot(data = titanic.train.clean,aes(x = Age)) +
  geom_histogram(aes(fill = as.factor(Survived)),binwidth = 2) +
  facet_wrap(~ Family)
```
Passanger who was travelling in family was survived better.

#### Divide data in training set and test set (730/161)
*Purpose to divide the training dataset to measure the efficiency of the model
```{r}
set.seed(2017)
library(rpart)
train.sample <- sample(891,730)
train.set <- titanic.train.clean[train.sample,]
test.set <- titanic.train.clean[-train.sample,]
```

#### Train a decision tree model on the data
```{r}
train.model <- rpart(Survived ~ Pclass + Sex + Age + Family + Fare + Embarked +Cabin,                        data = train.set, 
                       method = "class",
                       control = rpart.control(minsplit = 50, cp = 0))
```

#### Plot Decision Tree Model
```{r}
library(rpart.plot)
rpart.plot(train.model,digits = 3,fallen.leaves = TRUE,type = 3,extra = 101)

```

#### Evaluate Model Performance
```{r}
#test.set <- subset(test.set,select = c(-Survived))
pred.survived <- predict(train.model, newdata = test.set,type = "class")
```

```{r}
library(gmodels)
CrossTable(test.set$Survived,
           pred.survived,
           chisq = FALSE,
           prop.c = FALSE,
           prop.r = FALSE,
           dnn = c("Actual Value","Predicted Value"))
```
Efficiency of the model  = `(88+44)/161`

#### Result Prediction and saving output to a file
```{r}
pred <- predict(train.model, newdata = titanic.test.clean, type = "class")
```

*Save Result*
```{r}
solution <- data.frame(PassengerId = titanic.test.clean$PassengerId, 
                       Survived = pred)

write.csv(file = "solution.csv", x= solution)
```

**remove used variables**
```{r}
rm(list=ls())
gc()
```

