---
title: "ExploratoryDataAnalysis"
author: "Nilesh"
date: "8/2/2017"
output: html_document
---
```{r}
pf <- read.delim("https://s3.amazonaws.com/udacity-hosted-downloads/ud651/pseudo_facebook.tsv",sep  = '\t')
names(pf)
```

## Histogram of user's Birthday
```{r}
library(ggplot2)
ggplot(data = pf) + 
  geom_histogram(aes(dob_day)) +
  scale_x_continuous(breaks=1:33) +
  facet_wrap(~dob_month,nrow = 3)
```
### People born on 1st and 15th is almost2.5 times higher than any other day. This is quite strange. Intresting that default date 1 Jan has many fold as DOB. Seems it is default setting of FB or people don't want to share the personal information due to privacy.

## Understand outliers through Friend count
```{r}
ggplot(data = subset(pf,!is.na(gender))) + # Removing data for Gender is NA
  geom_histogram(aes(friend_count),binwidth = 25) +
  scale_x_continuous(limits = c(0,1000),breaks = seq(0,1000,50)) + # Limit the length of histogram
  facet_wrap(~gender)  # dividing histogram on gender basis to understand male/female, who have more friends

table(pf$gender)
by(pf$friend_count,pf$gender,summary)
```
## Examine the tenure data
```{r}
ggplot(data = pf) +
  geom_histogram(aes(tenure),color = "black",fill = "blue",binwidth = 30) # Monthly tenure data

ggplot(data = pf) +
  geom_histogram(aes(tenure/365),color = "black",fill = "orange",binwidth = 0.25) + # Changed to Yearly Tenure
  scale_x_continuous(breaks = seq(1,7,1),limits = c(0,7)) +
  xlab("No of Years using Facebook") +
  ylab("No of Users in Sample")
```
## Histogram based on age of the user
```{r}
ggplot(data = pf) +
  geom_histogram(aes(age),binwidth = 1,color = "red",fill="orange") +
  scale_x_continuous(breaks = seq(0,113,5),limits = c(0,113)) +
  xlab("Age of the User") +
  ylab("No of Users in sample")
```

## Transformation for Normal Distribution
```{r}
p1 <- ggplot(data = pf) + 
  geom_histogram(aes(friend_count))  # Actual data
p2 <- ggplot(data = pf) + 
  geom_histogram(aes(log10(friend_count ))) +  # transformed to log (added one to avoid infite)
  scale_x_log10()

p3<- ggplot(data = pf) + 
  geom_histogram(aes(sqrt(friend_count))) + # Square root transformation
  scale_x_sqrt()

library(gridExtra)
grid.arrange(p1,p2,p3,nrow = 1)

# other may, it may work
ggplot(data = pf) +
  geom_histogram(aes(friend_count)) +
  scale_x_log10()

ggplot(data = pf) +
  geom_histogram(aes(friend_count)) +
  scale_x_sqrt()
```

## Frequency Polygon( before we had histogram)
```{r warning=FALSE}
ggplot(data = subset(pf,!is.na(gender))) +
  geom_freqpoly(aes(friend_count,color=gender)) +
  scale_x_continuous(limits = c(0,1000),breaks = seq(0,1000,50))
```
# This plot doesn't say much about  who have more friend. Lets make it propotional polygon.
```{r}



```

```{r}
ggplot( data = subset(pf,!is.na(gender))) +
  geom_freqpoly(aes(x = friend_count, y = ..count../sum(..count..),color = gender))  +
  scale_x_continuous(breaks = seq(0,1000,50),limits = c(0,1000)) +
  xlab("Friend Counts") +
  ylab("Propotion of users with that friend count")


ggplot( data = subset(pf,!is.na(gender))) +
  geom_freqpoly(aes(x = friend_count, y = ..density..,color = gender))  +
  scale_x_continuous(breaks = seq(0,1000,50),limits = c(0,1000)) +
  xlab("Friend Counts") +
  ylab("Density Plot friend count")

```
## Determine which gender makes more likes on www_likes

```{r}
library(dplyr)
ggplot( data = filter(pf,!is.na(gender)) ) +
  geom_freqpoly(aes(x = www_likes , color = gender)) +
  scale_x_continuous(breaks = seq(0,50,10),limits = c(0,50))

```
## How many www_likes for make and femake

```{r}
pf %>% group_by(gender) %>% summarise(sum(www_likes))
```


## Boxplot
```{r}
ggplot( data = filter(pf,!is.na(gender))) +
  geom_boxplot(aes(x = gender, y = friend_count)) +
  xlab("Gender") +
  ylab("Friend Count") +
  coord_cartesian(ylim = c(0,300))

## Other way
ggplot( data = filter(pf,!is.na(gender))) +
  geom_boxplot(aes(x = gender, y = friend_count)) +
  xlab("Gender") +
  ylab("Friend Count") +
  scale_y_continuous(limits = c(0,300))  # difference but it removing data point, upper quartile value is different

group_by(pf,gender) %>% summarise(mean(friend_count))
```
## "Who made more friend requests on average - men or women?"
```{r}
ggplot(data = filter(pf,!is.na(gender))) +
  geom_boxplot(aes(x = gender , y = friendships_initiated )) +
  coord_cartesian(ylim = c(0,250))
summary(subset(pf,gender == "male")$friendships_initiated)
summary(subset(pf,gender == "female")$friendships_initiated)
by(pf$friendships_initiated,pf$gender,summary)
```
## Getting Logical
What percentange of Check in using Mobile?
```{r}
pf$mobile_check_in  <- NA
pf$mobile_check_in <- ifelse(pf$mobile_likes > 0, 1,0)
sum(pf$mobile_check_in)/length(pf$mobile_check_in)
```
