---
title: "Credit Data C5.0 Decision Tree"
output: html_notebook
---
**Include required Libraries**
```{r}
library(C50)
library(gmodels)
```

####Training C5.0 Model on training data
```{r}
model <- C5.0(train.set[-17],train.set$default)
summary(model)
```
#### Evaluate the performance of the model
```{r}
test.predict <- predict(model,test.set)
CrossTable(test.predict,test.set$default,
           dnn =c("Predicted Result","Actual Result"))

```

27% error, predicted crrect default 14 out of 33 which is just 42% correct prediction for defaults. 

#### Improve Model Performance
```{r}
model_boost_10 <- model <- C5.0(train.set[-17],train.set$default,trials = 10,rules = FALSE)
#summary(model)
test.predict_10 <- predict(model_boost_10,test.set)
CrossTable(test.set$default,test.predict_10,
           dnn =c("Actual Result","Predicted Result"),
           prop.r = FALSE,
           prop.c = FALSE,
           chisq = FALSE )
```

20/33 improvement from 14/33.

*Making mistake more costlier*

```{r}
matrix_diam <- list(c("no","yes"),c("no","yes"))
names(matrix_diam) = c("predicted","actual")
matrix_diam
error_cost = matrix(c(0,1,3,0),nrow = 2,dimnames  = matrix_diam)
error_cost
```
#### Add Error Cost
*This will make model to more conservative and predict more default to avoid load defaults.*
```{r}
model_boost_10 <- model <- C5.0(train.set[-17],
                                train.set$default,
                                trials = 10,
                                costs = error_cost)
#summary(model)
test.predict_10 <- predict(model_boost_10,test.set)
CrossTable(test.set$default,test.predict_10,
           dnn =c("Actual Result","Predicted Result"),
           prop.r = FALSE,
           prop.c = FALSE,
           chisq = FALSE )
```

